An Efficient Construction of the Prefer-Max Sequence and of an Infinite de Bruijn Sequence

### 2016-12-27 paper.tex
...

\title{An Infinite de Bruijn Sequence}

@@

While our main focus is on analyzing the reversed sequence, we also have a result about the prefer-max sequence itself. Our result gives the first (to our knowledge) efficient construction for that sequence:

\begin{algorithm}[!h]
	Append $n-1$ zeros and then the symbol $k$. Then, repeatedly, let $\sigma w$ be the word formed by the $n$ most recent symbols. Append the symbol $\sigma-1$  if $\sigma w$  is of the form specified in Algorithm~\ref{inf-construction}. Otherwise, if there is a maximal $\sigma'$ such that $\sigma'w$ is of that form, append the symbol $\sigma'$. Otherwise, if there is no such $\sigma'$, append the symbol $\sigma$.

@@

Each step of Algorithm~\ref{efficient-pref-max} can be computed in $O(n^2)$ time because it suffices to examine only $\sigma'$s which are symbols in $\sigma w$ or are a symbol that smaller by one than a symbol in $\sigma w$ (see Proposition~\ref{efficiency-of-pref-max} below).
The fact that this is indeed the prefer-max sequence is stated in the following theorem:

@@

The main tool that brought us to the above results is a formulation of an alternative description of the infinite sequence. For this, we need to introduce some language.

>>>>

\title{An Efficient Construction of the Prefer-Max Sequence and of an Infinite de Bruijn Sequence}

@@

While our main focus is on analyzing the reversed sequence, we also have a result about the prefer-max sequence itself. Our result gives the first (to the best of our knowledge) efficient construction for that sequence:

\begin{algorithm}
	Append $n-1$ zeros and then the symbol $k$. Then, repeatedly, let $\sigma w$ be the word formed by the $n$ most recent symbols. Append the symbol $\sigma-1$  if $\sigma w$  is of the form specified in Algorithm~\ref{inf-construction}. Otherwise, if there is a maximal $\sigma'$ such that $\sigma'w$ is of that form, append the symbol $\sigma'$. Otherwise, if there is no such $\sigma'$, append the symbol $\sigma$.

@@

Each step of Algorithm~\ref{efficient-pref-max} can be computed in $O(n^2)$ time because it suffices to examine for $\sigma'$ only symbols that are $\sigma w$ or are smaller by one than a symbol in $\sigma w$, as we will establish later in the paper.
The fact that this is indeed the prefer-max sequence is stated in the following theorem:

@@

A central tool to our analysis of the sequences descibed above is an alternative formulation of the Ford sequence. For this, we need to introduce some language, as follows.

The $n$ dimensional infinite de Bruijn graph is the graph $G_{n}=\T{V,E}$ where $V=\N^n$ and $E=\{ \T{\sigma_1 w, w \sigma_2} \colon \sigma_1,\sigma_2\in\N, w \in \N^{n-1}\}$. The de Bruijn sequences, defined in the beginning of this paper, are in one-to-one correspondence with the Hamiltonian cycles in the de Bruijn graph, as follows: (1) Concatenating the first letter of each vertex in a cycle produces a de Bruijn sequence of order $n$; and (2) A cycle can be constructed from a de Bruijn sequence $\T{\sigma_1,\dots,\sigma_{2^n}}$ of order $n$ by visiting the vertex $\sigma_1\cdots\sigma_n$, then $\sigma_2\cdots\sigma_{n+1}$ and so on.

Let $rtl(\sigma_1\sigma_2\dots\sigma_n) = \sigma_2\dots\sigma_n\sigma_1$ denote the rotation of a word to the left and let $rtl^i(w)$ denote a rotation of $i$ letters to the left. A \emph{necklace} in the de Bruijn graph $G_n$ is the set $\{rtl^i(w)\}_{i=1}^n$ where $w \in \N^n$, i.e., it is an equivalence class of $n$-character strings over $\N$ when strings that are rotations of each other are considered equivalent. A $w\in\N^n$ is called a \emph{key word} if it is larger in right-to-left lexicographic order than all of its rotations (all the words in its necklace).  Note the resemblance of this to the notion of Lyndon words~\cite{??}, on which we will elaborate later.

We are now ready to introduce another construction:

\begin{algorithm}[!h]
	Start with the one element sequence $\T{0^n}$. The sequence at the $i$th step is constructed by adding words to the previous sequence as follows: Let $k_i$ be the $i$th key word in right-to-left lexicographic order. Let $l$ be the number of leading zeros in $k_i$ ($l$ may be zero) and let $\sigma \in \N$ and $w\in\N^{n-l-1}$ be such that $k_i=0^l(\sigma+1)w$. Now, add the sub-sequence $\T{rtl^i(w0^l(\sigma+1))}_{i=0}^{n-1}$ between the word $\sigma w 0^l$ and its follower in the previously constructed sequence.
	\caption{A necklace joining construction.}
	\label{necklace-joining}
\end{algorithm}

To following proposition establishes that the above construction describes an infinite sequence of words:

\begin{proposition}
	For every $l>0$ there is $m>0$ such that for every $i>m$ the insertions of words at the $i$th step of Algorithm~\ref{necklace-joining} are all after the $l$th element of the sequence.
\end{proposition}

Let $S=\T{w_i}_{i=0}^\infty$ be the infinite sequence all whose prefixes agree with infinitely many prefixes of the finite sequences generated at the steps of Algorithm~\ref{necklace-joining}. Since every word is a rotation of some key word, we have that this sequence contains all the words in $\N^n$. Since there is a directed edge in $G_n$ from each word in this sequence to its follower, the sequence corresponds to an Hamiltonian path in $G_n$. The following theorem states that this path maps, by the standard correspondence described above, to the reverse of the prefer-map sequence:

\begin{theorem}
	If we take the first letter of each word in $S$ we get the sequence described in Algorithm~\ref{inf-construction}.
\end{theorem}

The
